Sensitivity Analysis

### 2023-08-24 SFE_article_Sensitivity_Analysis/Article_Model_Parameters.tex
24.08

% ===================================================
% METADATA
% ===================================================
\title[mode=title]{Parameter estimation}                      
\shorttitle{Parameter estimation}

@@

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
This study aimed to investigate the supercritical extraction process of the caraway oil from caraway seeds. The extraction was performed in a partially filled extractor with a fixed bed operated under multiple constant operating conditions. The dataset collected from four experiments consists of four experiments conducted at different operating conditions: $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar, and $50^\circ C$/$300$ bar. A distributed-parameter model is used to describe the fluid-solid extraction process with $CO_2$ as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, saturation concentration, and an initial state estimate. These parameters are estimated from the Peng-Robinson equation of state and by applying the maximum likelihood estimation method on yield data under the normal error assumption.

@@

\subsection{Parameter estimation} \label{CH: Parameter_estimation}
\subfile{Sections/Parameter_estimation}

\subsection{Experimental work}
\subfile{Sections/Experiments}

\section{Results}
\subfile{Sections/Results}

@@

This study presented basic aspects regarding the laboratory work for obtaining caraway extracts using supercritical $CO_2$ at different operating conditions. A first-principle process model was developed based on governing equations and simplified using the Low-Mach number approximation. Based on the literature review, the extraction kinetic model was selected. The values of unknown parameters in the process model were obtained by parameter estimation. The yield data from the experiments were compared against the yield data generated by the model. The IPOPT solver was used to solve the maximum likelihood estimation. The process model was found to have good agreement with experimental yield curves. Later, the parameters found for each experiment were combined to obtain more general relationships. The correlations for ${\color{orange}D_i^R}$, ${\color{orange}\Upsilon}$ and ${\color{orange}D_e^M}$ are presented as a function of ${\color{orange}\rho}$ or ${\color{orange}R_e}$. It should be noted that obtained correlations have been prepared with a limited amount of data. These correlations can be used to study qualitatively the impact of different operating conditions on the extraction yield by different methods, such as sensitivity analysis. Moreover, the generalized model could be useful in finding the optimal operating conditions with respect to the process economy.

@@

\subsection{Maximum Likelihood} \label{CH: ML}
\subfile{Sections/Likelihood}

@@

\subsection{Concentration profiles} \label{CH: Profiles}
\subfile{Sections/Profiles}

>>>>

% ===================================================
% METADATA
% ===================================================
\title[mode=title]{Sensitivity Analysis}                      
\shorttitle{Sensitivity Analysis}

@@

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study aimed to investigate the supercritical extraction process of caraway oil from caraway seeds. The extraction was performed in a partially filled extractor with a fixed bed operated under multiple operating conditions. A distributed-parameter model describes the fluid-solid extraction process with $CO_2$ as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The local sensitivity analysis investigates the influence of infinitely small changes in the inlet temperature, pressure, and flow rate on the extraction yield.

@@

\subsection{Sensitivity Analysis} \label{CH: Sensitivity_Analysis}
\subfile{Sections/Sensitivity_Analysis}

%\subsection{Experimental work}
%\subfile{Sections/Experiments}

\section{Results}
\subfile{Sections/Results_Sensitivity}

@@

@@

%\subsection{Maximum Likelihood} \label{CH: ML}
%\subfile{Sections/Likelihood}

@@

%\subsection{Concentration profiles} \label{CH: Profiles}
%\subfile{Sections/Profiles}

### 2023-08-24 SFE_article_Sensitivity_Analysis/Sections/Model.tex
24.08

	\subsubsection{Continuity equation} \label{CH: Continuity}
	 %When the cross-sectional area of the channel $A_f$ is specified as a function of the void fraction ${\color{magenta}\phi}(z)$ (where ${\color{magenta}\phi}$ is the void fraction of the bed and $A$ is the cross-section of the empty extractor), the continuity equation takes the form: %\todo[]{assign a color of v and u}
	The obtained above quasi-one-dimensional continuity equation (Equation \ref{EQ: CompressibleEuler_1}) is further modified by specifying a function ${\color{magenta}A_f}(z) = {\color{magenta}A_f\phi}(z)$ to take into account the change of the cross-section available for the fluid. The cross-section change is taken into account by void fraction ${\color{magenta}\phi}(z)$, which depends on the spatial dimension. The equation \ref{EQ: Continuity_differential} shows the differential form, of the continuity equation:

@@

	{\color{red}assign a color of v and u}

@@

	The detailed derivation of the mass balance equation for the fluid phase can be found in the appendix (\ref{CH: Gouverning equations}). The movement of the mobile pseudo-homogeneous phase (Equation \ref{Model_fluid}) is considered only in the axial direction, while the properties of the system in the radial direction are assumed to be uniform. Additionally, the boundary layer adjacent to the inner wall of the extractor is neglected, resulting in a constant velocity profile across any cross-section of the extractor perpendicular to the axial direction. Although the particle size distribution and void fraction of the solid phase may change along the extractor, they are assumed to remain constant in time. Furthermore, the thermodynamic pressure is assumed to be constant along the device due to the Low-Mach number condition, as previously discussed. The amount of solute in the solvent is considered negligible, resulting in the fluid phase being described as pseudo-homogeneous, and its properties are assumed to be the same as the solvent. The mass balance equation for the fluid phase includes convection, diffusion, and kinetic terms.

@@

	As the solvent flows through the bed, $CO_2$ molecules diffuse into the pores and adsorb on the particle surface to form an external fluid film around the solid particles due to the solvent-solid matrix interactions. The effect of Knudsen diffusion is negligible in this process, as the mean free path of the molecule is much smaller than the pore diameter. The dissolved solute diffuses from the particle's core through the solid-fluid interface, the pore, and the film into the bulk. Figure \ref{fig: SFE_Mechanism} illustrates the mass transfer mechanism, where the mean solute concentration in the solid phase is denoted as ${\color{blue}c_s}$ and the equilibrium concentrations at the solid-fluid interface are denoted as ${\color{blue}c_s^*}$ and ${\color{blue}c_P^*}$, respectively, for solid and fluid phases. The concentration of the solutes in the fluid phase in the center of the pore is denoted as ${\color{blue}c_{P}}$. As the solute diffuses through the pore, its concentration changes and reaches ${\color{blue}c_{Pf}}$ at the opening of the pore. The solute then diffuses through the film around the particle and reaches a concentration in bulk ${\color{blue}c_f}$. The two-film theory describes the solid-fluid interface inside the pore. The overall mass transfer coefficient can be determined if the relationship between the solute concentration in one phase and its equilibrium concentration is known.

		\begin{figure}[h!]
			\centering
			%	\includegraphics[trim  = 6cm 0cm 3cm 0cm,clip,width=\linewidth]{Figures/SFE_draft.png}
			\includegraphics[trim = 5.8cm 1.1cm 6cm 3cm,clip,width=\columnwidth]{Figures/SFE_draft.pdf}	
			\caption{The extraction mechanism}
			\label{fig: SFE_Mechanism}
		\end{figure}

		\citet{Bulley1984} suggests a process where the driving force for extraction is given by the difference between the concentration of the solute in bulk, ${\color{blue}c_f}$, and in the center of the pore, ${\color{blue}c_P^*}$. The concentration ${\color{blue}c_P^*}$ is in equilibrium with ${\color{blue}c_s}$ according to an equilibrium relationship. The rate of extraction is thus ${\color{blue}r_\text{e}}\left({\color{blue}c_f} - {\color{blue}c^*_P}({\color{blue}c_s})\right)$.

		On the other hand, \citet{Reverchon1996} proposes a driving force given by the difference between ${\color{blue}c_s}$ and ${\color{blue}c_P^*}$. ${\color{blue}c_P^*}$ is determined by an equilibrium relationship with ${\color{blue}c_f}$ and the extraction rate is ${\color{blue}r_\text{e}}\left({\color{blue}c_s} - {\color{blue}c^*_P}({\color{blue}c_f})\right)$ or more precisely

@@

			The Equation \ref{EQ:Enthalpy_PR} requires an reference sate, which in this case is assumed to be $T_{ref}=298.15~[K]$ and $P_{ref}=1.01325~[bar]$.

@@

		Such a simplified equation takes into account the pressure change in the energy balance, but the dynamics are simplified and do not consider the effects of pressure losses. In a real system, the pressure change dynamics would depend on the pump used in the system, as well as the back-pressure regulator used to control the outlet valve.

@@

		The efficiency of the process (the yield) is calculated according to equations \ref{Model_measurment_1} as presented by \citet{Sovova1994a}. The measurement equation evaluate the mass of solute at the outlet of the extraction unit and sums it. The integral form of the measurement equation (\ref{Model_measurment_1}) can be transformed into the differential form (\ref{Model_measurment_2}) and augmented with the process model.

@@

		\subsubsection{Initial and boundary conditions} 
		It is assumed that the solvent is free of solute at the entrance of the extractor and that all the solid particles have the same initial solute content ${\color{magenta}c_{s0}}$. On the other hand, the solute concentration in the fluid phase should not follow the same assumption as the solid phase. During the preparation period, the solute diffuses to the fluid phase in contact with the solid particles. Later, the solute in the fluid phase is partially moved along the extractor (if the pressure increase in the system, the pump cause the movement of the fluid, even if the outlet valve is closed). As a result, the distribution of solute mass in the fluid phase is assumed not to be uniform. Some conclusions can be drawn from the analysis of the initial part of each yield curve obtained from the laboratory (Figure \ref{fig: estimation_results}). It can be noticed that each curve at the beginning has a curvature, which is not linear. Generally, a quadratic function could approximate the initial part of each extraction curve. A function that, after integration, gives a quadratic-like result is a straight line. Based on that observation, the solute concentration in the fluid phase is assumed to be linearly distributed. The solute concentration is assumed to be zero at the outlet and reach the maximum at the beginning of the fixed bed. The details on the calculation are given in Appendix \ref{CH: IC_BC}. The linear distribution $H(z)$ can be defined if the total mass of solute $m_{total}$ and initial mass ratio between solid and fluid phases $\tau$ are known.
		Moreover, it is considered that the initial temperature of the extractor in every place is the same and described by ${\color{magenta}h_0}$. Therefore, the initial conditions employed in the simulation are:

@@

			In a state-space sense, the state variables of the system are the local concentrations of solute in the fluid and solid phases (${\color{blue}c_f}(t,z)$ and ${\color{blue}c_s}(t,z)$, respectively), and the local enthalpy of the pseudo-homogeneous phase (${\color{blue}h}(t,z)$). The controllable input variables are the mass flow-rate and temperature of the solvent in the feed and the pressure in the extractor. We extend the system state-space by assuming that extraction yield can be modelled as a function of a known initial mass of solute in the solid phase and it can be measured after the separator (${\color{blue}Y}(t)$). The system is controllable by manipulating the flow-rate and temperature (enthalpy) of $CO_2$ in the feed, and the pressure in the extractor.

>>>>

	\subsubsection{Continuity equation} \label{CH: Continuity}
	 %When the cross-sectional area of the channel $A_f$ is specified as a function of the void fraction ${\color{magenta}\phi}(z)$ (where ${\color{magenta}\phi}$ is the void fraction of the bed and $A$ is the cross-section of the empty extractor), the continuity equation takes the form: %\todo[]{assign a color of v and u}
	The obtained above quasi-one-dimensional continuity equation (Equation \ref{EQ: CompressibleEuler_1}) is further modified by specifying a function ${\color{magenta}A_f}(z) = {\color{magenta}A_f\phi}(z)$ to take into account the change of the cross-section available for the fluid. The cross-section change is taken into account by void fraction ${\color{magenta}\phi}(z)$, which depends on the spatial dimension. The Equation \ref{EQ: Continuity_differential} shows the differential form of the continuity equation:

@@

	%{\color{red}assign a color of v and u}

@@

	The detailed derivation of the mass balance equation for the fluid phase can be found in the appendix (\ref{CH: Gouverning equations}). The movement of the pseudo-homogeneous fluid phase (Equation \ref{Model_fluid}) is considered only in the axial direction, while the properties of the system in the radial direction are assumed to be uniform. Additionally, the boundary layer adjacent to the inner wall of the extractor is neglected, resulting in a constant velocity profile across any cross-section of the extractor perpendicular to the axial direction. Although the particle size distribution and void fraction of the solid phase may change along the extractor, they are assumed to remain constant in time. Furthermore, the thermodynamic pressure is assumed to be constant along the device due to the Low-Mach number condition, as previously discussed. The amount of solute in the solvent is considered negligible, resulting in the fluid phase being described as pseudo-homogeneous, and its properties are assumed to be the same as the solvent. The mass balance equation for the fluid phase includes convection, diffusion, and kinetic terms.

@@

	As the solvent flows through the bed, $CO_2$ molecules diffuse into the pores and adsorb on the particle surface to form an external fluid film around the solid particles due to the solvent-solid matrix interactions. The effect of Knudsen diffusion is negligible in this process, as the mean free path of the molecule is much smaller than the pore diameter. The dissolved solute diffuses from the particle's core through the solid-fluid interface, the pore, and the film into the bulk. Figure \ref{fig: SFE_Mechanism} illustrates the mass transfer mechanism, where the mean solute concentration in the solid phase is denoted as ${\color{blue}c_s}$ and the equilibrium concentrations at the solid-fluid interface are denoted as ${\color{blue}c_s^*}$ and ${\color{blue}c_p^*}$, respectively, for solid and fluid phases. The concentration of the solutes in the fluid phase in the center of the pore is denoted as ${\color{blue}c_{p}}$. As the solute diffuses through the pore, its concentration changes and reaches ${\color{blue}c_{pf}}$ at the opening of the pore. The solute then diffuses through the film around the particle and reaches a concentration in bulk ${\color{blue}c_f}$. The two-film theory describes the solid-fluid interface inside the pore. The overall mass transfer coefficient can be determined if the relationship between the solute concentration in one phase and its equilibrium concentration is known.

		\begin{figure}[h!]
			\centering
			\includegraphics[trim = 45cm 0cm 60cm 20cm,clip,width=\columnwidth]{Figures/SFE_PFD.drawio.png}	
			\caption{The extraction mechanism}
			\label{fig: SFE_Mechanism}
		\end{figure}

		\citet{Bulley1984} suggests a process where the driving force for extraction is given by the difference between the concentration of the solute in bulk, ${\color{blue}c_f}$, and in the center of the pore, ${\color{blue}c_p^*}$. The concentration ${\color{blue}c_p^*}$ is in equilibrium with ${\color{blue}c_s}$ according to an equilibrium relationship. The rate of extraction is thus ${\color{blue}r_\text{e}}\left({\color{blue}c_f} - {\color{blue}c^*_p}({\color{blue}c_s})\right)$.

		On the other hand, \citet{Reverchon1996} proposes a driving force given by the difference between ${\color{blue}c_s}$ and ${\color{blue}c_p^*}$. ${\color{blue}c_p^*}$ is determined by an equilibrium relationship with ${\color{blue}c_f}$ and the extraction rate is ${\color{blue}r_\text{e}}\left({\color{blue}c_s} - {\color{blue}c^*_p}({\color{blue}c_f})\right)$ or more precisely

@@

			The Equation \ref{EQ:Enthalpy_PR} requires an reference sate, which in this case is assumed to be ${\color{magenta}T_{ref}}=298.15~[K]$ and ${\color{magenta}P_{ref}}=1.01325~[bar]$.

@@

		Such a simplified equation takes into account the pressure change in the energy balance, but the dynamics are simplified and do not consider the effects of pressure losses. In a real system, the dynamics of pressure change would depend on a pump used in an extraction system, as well as a back-pressure regulator used to control an outlet valve.

@@

		The efficiency of the process (the yield) is calculated according to Equation \ref{Model_measurment_1} as presented by \citet{Sovova1994a}. The measurement equation evaluate the mass of solute at the outlet of the extraction unit and sums it. The integral form of the measurement equation (\ref{Model_measurment_1}) can be transformed into the differential form (\ref{Model_measurment_2}) and augmented with the process model.

@@

		\subsubsection{Initial and boundary conditions} 
		It is assumed that the solvent is free of solute at the entrance of the extractor and that all the solid particles have the same initial solute content ${\color{magenta}c_{s0}}$. On the other hand, the solute concentration in the fluid phase should not follow the same assumption as the solid phase. Every SFE system needs some time to reach the desired operating conditions but the solute diffuses from solid phase to the fluid phase already at the preparation stage. For instance, a pump introduce more and more fluid to the extractor to increase the pressure, which makes the fluid present already in that vessel to moves internally hence the solute in the fluid phase is non-uniformly distributed. Some conclusions can be drawn from the analysis of the initial part of each yield curve obtained from the laboratory (Figure \ref{fig: estimation_results}). It can be noticed that each curve at the beginning has a curvature, which is not linear. Generally, a quadratic function could approximate the initial part of each extraction curve. A function that, after integration, gives a quadratic-like result is a straight line. Based on that observation, the solute concentration in the fluid phase is assumed to be linearly distributed. The solute concentration is assumed to be zero at the outlet and reach the maximum at the beginning of the fixed bed. The details on the calculation are given in Appendix \ref{CH: IC_BC}. The linear distribution $H(z)$ can be defined if the total mass of solute $m_{total}$ and initial mass ratio between solid and fluid phases $\tau$ are known.
		Moreover, it is considered that the initial temperature of the extractor in every place is the same and described by ${\color{magenta}h_0}$. Therefore, the initial conditions employed in the simulation are:

@@

			In a state-space sense, the state variables of the system are the local concentrations of solute in the fluid and solid phases (${\color{blue}c_f}(t,z)$ and ${\color{blue}c_s}(t,z)$, respectively), and the local enthalpy of the pseudo-homogeneous phase (${\color{blue}h}(t,z)$). The controllable input variables are the mass flow-rate and temperature of the solvent in the feed and the pressure in the extractor. Additionally, the pressure change is augumented with the state-space and denoted as ${\color{blue}P}(t)$.
			The system state-space is extended by assuming that extraction yield can be modelled as a function of a known initial mass of solute in the solid phase and it can be measured after the separator (${\color{blue}Y}(t)$). The system is controllable by manipulating the flow-rate and temperature (enthalpy) of $CO_2$ in the feed, and the pressure in the extractor.

### 2023-08-24 SFE_article_Sensitivity_Analysis/Sections/Parameter_estimation.tex
24.08

	The goal of parameter estimation is to obtain the "best" estimate of parameter set ${\color{magenta} \theta}$ (which is a subset of the parameter space ${\color{magenta}\Theta}$ containing all parameters of a model) based on the continuous observations ${\color{blue}Y}(t)$ or the discrete observations ${\color{blue}Y}(t_i)$. Conceptually, the unobservable error $\epsilon(t)$ is added to the deterministic model output, ${\color{blue}y}(t)$ (Equation \ref{Model_measurment_1}), to give the observable dependent variable ${\color{blue}Y}(t)$ (for example results of an experiment). For discrete observations, this can be expressed as:

>>>>

	Not all the parameters present in a process model can be estimated from the theoretical considerations. The goal of parameter estimation is to obtain the "best" estimate of unknown parameters ${\color{magenta} \theta}$ (which is a subset of the parameter space ${\color{magenta}\Theta}$ containing all parameters of a model) based on the continuous observations ${\color{blue}Y}(t)$ or the discrete observations ${\color{blue}Y}(t_i)$. Conceptually, the unobservable error $\epsilon(t)$ is added to the deterministic model output, ${\color{blue}y}(t)$ (Equation \ref{Model_measurment_1}), to give the observable dependent variable ${\color{blue}Y}(t)$ (for example results of an experiment). For discrete observations, this can be expressed as:

### 2023-08-24 SFE_article_Sensitivity_Analysis/Sections/Thermodynamic_details.tex
24.08

	The specific heat $C_p^{\text{F}}$ can be calculated from the equation of state, again under the assumption that the fluid phase consists of pure carbon dioxide and that the specific heat of real fluids can be calculated from an ideal contribution plus a residual term \citet{Pratt2001}. The main steps in the derivation of ${\color{orange}C^{CO_2}_p}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$ are presented below:

	{\footnotesize
	\begin{subequations}\begin{align}
			{\color{orange}C^{CO_2}_v}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right) & = {\color{orange}C_v^{\text{I}}}\left({\color{blue}T}(t,z),P(t)\right) + {\color{orange}C_v^{\text{R}}}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right) \label{eq: Cv_CO2}; \\ 
			{\color{orange}C^{CO_2}_p}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right) & = \underbrace{{\color{orange}C_p^{\text{I}}}\left({\color{blue}T}(t,z),P(t)\right)}_{\text{Eq. } \eqref{eq: Cp_I}} + \underbrace{{\color{orange}C_p^{\text{R}}}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)}_{\text{Eq. } \eqref{eq: Cp_R}}  \label{eq: Cp_CO2}.
	\end{align}\end{subequations} }

	${\color{orange}C^{CO_2}_v}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$ and ${\color{orange}C^{CO_2}_p}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$ are the specific heat of $CO_2$ at constant volume and pressure, respectively. ${\color{orange}C_v^{\text{I}}}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$ and ${\color{orange}C_p^{\text{I}}}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$, with ${\color{orange}C_p^{\text{I}}}({\color{blue}T}(t,z)) - {\color{orange}C_v^{\text{I}}}({\color{blue}T}(t,z)) = {\color{magenta}R}$, are the specific heat of an ideal gas at constant volume and pressure. ${\color{orange}C_v^{\text{R}}}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$ and ${\color{orange}C_p^{\text{R}}}({\color{blue}T}(t,z), {\color{red}P}(t))$ are the correction terms. \\

	For $CO_2$, the ideal gas contribution to the specific heat at constant ${\color{red}P}(t)$, as function of ${\color{blue}T}(t,z)$ is given by,

@@

	where the coefficients ({\color{red}REF}) of the expansion are ${\color{magenta}C_{P0}} = 4.728$, ${\color{magenta}C_{P1}} = 1.75 \times 10^{-3}$, ${\color{magenta}C_{P2}} = -1.34 \times 10^{-5}$, and ${\color{magenta}C_{P3}} = 4.10 \times 10^{-9}$. For the correction term ${\color{orange}C_p^{\text{R}}}\left({\color{blue}T}(t,z){\color{red}P}(t)\right)$ at constant pressure ${\color{red}P}(t)$, we have

@@

	For the partial derivative of the volume with respect to temperature $T$ at constant pressure $P(t)$, we have

@@

	with partial derivative of the compressibility factor with respect to temperature ${\color{blue}T}$ at constant pressure ${\color{blue}P}(t)$

@@

	{\footnotesize
	\begin{equation}
		{\color{orange}C_v^{\text{R}}}\left({\color{blue}T}(t,z),{\color{red}P}(t)\right) = \left(\cfrac{\partial{\color{orange}U^{\text{R}}}\left({\color{blue}T}(t,z),{\color{red}P}(t)\right)}{\partial {\color{blue}T}}\right)_{{\color{orange}v_m}(t,z)}. \label{eq: CvR}
	\end{equation} }

>>>>

	%The specific heat $C_p^{\text{F}}$ can be calculated from the equation of state, again under the assumption that the fluid phase consists of pure carbon dioxide and that the specific heat of real fluids can be calculated from an ideal contribution plus a residual term \citet{Pratt2001}. The main steps in the derivation of ${\color{orange}C_p}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$ are presented below: 
	Following the work of \citet{Pratt2001}, it is assumed that the specific heat can be obtained as a sum of ideal gas contribution and a residual correction for non-ideal behaviour, as given by Equations \ref{eq: Cv_CO2} and \ref{eq: Cp_CO2}.

	{\footnotesize
	\begin{subequations}\begin{align}
			{\color{orange}C_v}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right) & = {\color{orange}C_v^{\text{I}}}\left({\color{blue}T}(t,z),P(t)\right) + {\color{orange}C_v^{\text{R}}}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right) \label{eq: Cv_CO2}; \\ 
			{\color{orange}C_p}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right) & = \underbrace{{\color{orange}C_p^{\text{I}}}\left({\color{blue}T}(t,z),P(t)\right)}_{\text{Eq. } \eqref{eq: Cp_I}} + \underbrace{{\color{orange}C_p^{\text{R}}}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)}_{\text{Eq. } \eqref{eq: Cp_R}}  \label{eq: Cp_CO2}.
	\end{align}\end{subequations} }

	${\color{orange}C_v}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$ and ${\color{orange}C_p}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$ are the specific heat of $CO_2$ at constant volume and pressure, respectively.

	The ideal-gas contribution ${\color{orange}C_v^{\text{I}}}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$ and ${\color{orange}C_p^{\text{I}}}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)$, can be obtained based on heat-capacity data applicable to gases at low pressure. For $CO_2$, the ideal gas contribution to the specific heat at constant ${\color{red}P}(t)$, as function of ${\color{blue}T}(t,z)$ is given by,

@@

	where the coefficients ({\color{red}REF}) of the expansion are ${\color{magenta}C_{P0}} = 4.728$, ${\color{magenta}C_{P1}} = 1.75 \times 10^{-3}$, ${\color{magenta}C_{P2}} = -1.34 \times 10^{-5}$, and ${\color{magenta}C_{P3}} = 4.10 \times 10^{-9}$.

	The correction term ${\color{orange}C_p^{\text{R}}}\left({\color{blue}T}(t,z){\color{red}P}(t)\right)$ at constant pressure ${\color{red}P}(t)$, can be defined by introducing  ${\color{orange}C_p^{\text{I}}}({\color{blue}T}(t,z)) - {\color{orange}C_v^{\text{I}}}({\color{blue}T}(t,z)) = {\color{magenta}R}$ and using the definition of ${\color{orange}C_v^{\text{R}}}\left({\color{blue}T}(t,z){\color{red}P}(t)\right)$. The general form of the correction term is given by Equation \ref{eq: Cp_R}

@@

	For the partial derivative of the volume with respect to temperature ${\color{blue}T}(t,z)$ at constant pressure ${\color{red}P}(t)$, we have

@@

	with partial derivative of the compressibility factor with respect to temperature ${\color{blue}T}(t,z)$ at constant pressure ${\color{blue}P}(t)$

@@

	{\footnotesize
	\begin{equation}
		{\color{orange}C_v^{\text{R}}}\left({\color{blue}T}(t,z),{\color{red}P}(t)\right) = \left(\cfrac{\partial{\color{orange}U^{\text{R}}}\left({\color{blue}T}(t,z),{\color{red}P}(t)\right)}{\partial {\color{blue}T}}\right)_{{\color{orange}v_m}(t,z)} \label{eq: CvR}
	\end{equation} }
